---
title: "Business Intelligence Project"
author: "<Specify your name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | *\<list one student name, group, and ID per line; you should be between 2 and 5 members per group\>* |
|                                                   |                                                                                                      |
|                                                   | 1.  135232 - Starfield  - Sadiki Hamisi                                                              |
|                                                   |                                                                                                      |
|                                                   | 2. 134782 - Starfield - Yasmin Choma                                                                 |
|                                                   |                                                                                                      |
|                                                   | 3.  134783 - Starfield - Moses mbugua                                                                |
|                                                   |                                                                                                      |
|                                                   | 4.  122998 - Starfield - Glenn Oloo                             
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   |  Starfield                                                      
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                              |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                             |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                          |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                           |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

# Understanding the Dataset (Exploratory Data Analysis (EDA))

## Loading the Dataset

### Source:

The dataset that was used can be downloaded here: *\<provide a link\>*

### Reference:

*\<Cite the dataset here using APA\>\
Refer to the APA 7th edition manual for rules on how to cite datasets: <https://apastyle.apa.org/style-grammar-guidelines/references/examples/data-set-references>*


```{r Dataset Loader}

library(readr)
library(dplyr)
student_performance_dataset <-
  read_csv("data/student_performance_dataset.csv",
           col_types =
             cols(
               class_group = col_factor(levels = c("A", "B", "C")),
               gender = col_factor(levels = c("1", "0")),
               YOB = col_date(format = "%Y"),
               regret_choosing_bi = col_factor(levels = c("1", "0")),
               drop_bi_now = col_factor(levels = c("1", "0")),
               motivator = col_factor(levels = c("1", "0")),
               read_content_before_lecture =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               anticipate_test_questions =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               answer_rhetorical_questions =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               find_terms_I_do_not_know =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               copy_new_terms_in_reading_notebook =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               take_quizzes_and_use_results =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               reorganise_course_outline =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               write_down_important_points =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               space_out_revision =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               studying_in_study_group =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               schedule_appointments =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               goal_oriented = col_factor(levels = c("1", "0")),
               spaced_repetition =
                 col_factor(levels = c("1", "2", "3", "4")),
               testing_and_active_recall =
                 col_factor(levels = c("1", "2", "3", "4")),
               interleaving = col_factor(levels = c("1", "2", "3", "4")),
               categorizing = col_factor(levels = c("1", "2", "3", "4")),
               retrospective_timetable =
                 col_factor(levels = c("1", "2", "3", "4")),
               cornell_notes = col_factor(levels = c("1", "2", "3", "4")),
               sq3r = col_factor(levels = c("1", "2", "3", "4")),
               commute = col_factor(levels = c("1", "2", "3", "4")),
               study_time = col_factor(levels = c("1", "2", "3", "4")),
               repeats_since_Y1 = col_integer(),
               paid_tuition = col_factor(levels = c("0", "1")),
               free_tuition = col_factor(levels = c("0", "1")),
               extra_curricular = col_factor(levels = c("0", "1")),
               sports_extra_curricular = col_factor(levels = c("0", "1")),
               exercise_per_week = col_factor(levels = c("0", "1", "2", "3")),
               meditate = col_factor(levels = c("0", "1", "2", "3")),
               pray = col_factor(levels = c("0", "1", "2", "3")),
               internet = col_factor(levels = c("0", "1")),
               laptop = col_factor(levels = c("0", "1")),
               family_relationships =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               friendships = col_factor(levels = c("1", "2", "3", "4", "5")),
               romantic_relationships =
                 col_factor(levels = c("0", "1", "2", "3", "4")),
               spiritual_wellnes =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               financial_wellness =
                 col_factor(levels = c("1", "2", "3", "4", "5")),
               health = col_factor(levels = c("1", "2", "3", "4", "5")),
               day_out = col_factor(levels = c("0", "1", "2", "3")),
               night_out = col_factor(levels = c("0", "1", "2", "3")),
               alcohol_or_narcotics =
                 col_factor(levels = c("0", "1", "2", "3")),
               mentor = col_factor(levels = c("0", "1")),
               mentor_meetings = col_factor(levels = c("0", "1", "2", "3")),
               `Attendance Waiver Granted: 1 = Yes, 0 = No` =
                 col_factor(levels = c("0", "1")),
               GRADE = col_factor(levels = c("A", "B", "C", "D", "E"))),
           locale = locale())

View(student_performance_dataset)
```


```{Student Performance Dataset Dimensions}

dim(student_performance_dataset)
```


```{Student Performance Dataset Data Types}
sapply(student_performance_dataset, class)
glimpse(student_performance_dataset)

```


```{The Scale Basic Transform on the Student Performance Dataset}

# BEFORE
View(student_performance_dataset)

summary(student_performance_dataset)
student_performance_dataset_enj <- as.numeric(unlist(student_performance_dataset[, 49]))
hist(student_performance_dataset_enj, main = names(student_performance_dataset)[49])

model_of_the_transform <- preProcess(student_performance_dataset, method = c("scale"))
print(model_of_the_transform)
student_data_scale_transform <- predict(model_of_the_transform, student_performance_dataset)

# AFTER
summary(student_data_scale_transform)
student_performance_dataset_enj <- as.numeric(unlist(student_data_scale_transform[, 49]))
hist(student_performance_dataset_enj, main = names(student_data_scale_transform)[49])

```


```{Histogram vizualization before and after Transformation}

#BEFORE
hist(student_performance_dataset_enj, main = names(student_performance_dataset)[49])


#AFTER
hist(student_performance_dataset_enj, main = names(student_data_scale_transform)[49])
```


```{The Centre Basic Transform on the Student Performance Dataset}

summary(student_performance_dataset)
model_of_the_transform <- preProcess(student_performance_dataset, method = c("scale"))
print(model_of_the_transform)
student_data_scale_transform <- predict(model_of_the_transform, student_performance_dataset)
summary(student_data_scale_transform)

```


```{The Standardize Basic Transform on the Crop Dataset}

#BEFORE
summary(student_performance_dataset)
sapply(student_performance_dataset[, 49], sd)
model_of_the_transform <- preProcess(student_performance_dataset,
                                     method = c("scale", "center"))
print(model_of_the_transform)
student_performance_standardize_transform <- predict(model_of_the_transform, student_performance_dataset) # nolint

# AFTER
summary(student_performance_standardize_transform)
sapply(student_performance_standardize_transform[, 49], sd)

```


```{The Normalize Transform on the Student Performance Dataset}
summary(student_performance_dataset)
model_of_the_transform <- preProcess(student_performance_dataset, method = c("range"))
print(model_of_the_transform)
student_performance_normalize_transform <- predict(model_of_the_transform, student_performance_dataset)
summary(student_performance_normalize_transform)

```


```{Box-Cox Power Transform on the Student Performance Dataset}

# BEFORE
summary(student_performance_standardize_transform)

# Calculate the skewness before the Box-Cox transform
sapply(student_performance_standardize_transform[, 49],  skewness, type = 2)
sapply(student_performance_standardize_transform[, 49], sd)

model_of_the_transform <- preProcess(student_performance_standardize_transform,
                                     method = c("BoxCox"))
print(model_of_the_transform)
student_performance_box_cox_transform <- predict(model_of_the_transform,
                                                 student_performance_standardize_transform)

# AFTER
summary(student_performance_box_cox_transform)

sapply(student_performance_box_cox_transform[, 49],  skewness, type = 2)
sapply(student_performance_box_cox_transform[, 49], sd)


# Calculate the skewness after the Box-Cox transform
sapply(student_performance_box_cox_transform[, 49],  skewness, type = 2)
sapply(student_performance_box_cox_transform[, 49], sd)

```


```{Yeo-Johnson Power Transform on the Student Performance Dataset}
# BEFORE
summary(student_performance_standardize_transform)

# Calculate the skewness before the Yeo-Johnson transform
sapply(student_performance_standardize_transform[, 49],  skewness, type = 2)
sapply(student_performance_standardize_transform[, 49], sd)

model_of_the_transform <- preProcess(student_performance_standardize_transform,
                                     method = c("YeoJohnson"))
print(model_of_the_transform)
student_performance_yeo_johnson_transform <- predict(model_of_the_transform, # nolint
                                           student_performance_standardize_transform)

# AFTER
summary(student_performance_yeo_johnson_transform)

```


```{PCA for Dimensionality Reduction on the student perfomance dataset}

summary(student_performance_dataset)

model_of_the_transform <- preProcess(student_performance_dataset,
                                     method = c("scale", "center", "pca"))
print(model_of_the_transform)
student_performance_dataset_pca_dr <- predict(model_of_the_transform, student_performance_dataset)

summary(student_performance_dataset_pca_dr)
dim(student_performance_dataset_pca_dr)

```

